Causal Bandits: Learning Good Interventions via Causal Inference

### 2016-02-02 causal-bandit/icml-2015/introduction.tex

We focus on the simple regret, which measures the difference between the return of the optimal action and that of the action chosen by the algorithm after $T$ rounds.
In \S\ref{sec:simple-regret} we analyse a specific family of causal bandit problems that we call \emph{parallel bandit} problems in which $N$ factors affect the reward independently.
We propose a simple causal best arm identification algorithm (Algorithm~\ref{alg:simple}) for this problem and show that up to logarithmic factors it enjoys minimax optimal
simple regret guarantees of $\tilde\Theta(\sqrt{m/T})$ where $m$ depends on the causal model and may be much smaller than $N$.
In contrast, existing best arm identification algorithms suffer $\Omega(\sqrt{N/T})$ simple regret.
This shows theoretically the value of our framework over the traditional bandit problem. 
Experiments in \S\ref{sec:experiments} further demonstrate the value of causal models in bandit problems.

Interventions and observations may have a complex relationship in the general causal model. 
In \S\ref{sec:simple-regret-general} we propose an algorithm inspired by importance-sampling that (a) enjoys sub-linear regret equivalent to the optimal rate in the parallel 
bandit setting and (b) captures many (all?) of the intricacies of sharing information in a causal graph in the general case. As in the parallel bandit case, the regret guarantee
is often significantly better than $O(\sqrt{|\calA|/T})$ and never worse, where $|\calA|$ is the number of available interventions.

>>>>

We focus on the simple regret, which measures the difference between the return of the optimal action and that of the action chosen by the algorithm after $T$ rounds.
In \S\ref{sec:simple-regret} we analyse a specific family of causal bandit problems that we call \emph{parallel bandit} problems in which $N$ factors affect the reward independently and there are $2N$ possible interventions.
We propose a simple causal best arm identification algorithm (Algorithm~\ref{alg:simple}) for this problem and show that up to logarithmic factors it enjoys minimax optimal
simple regret guarantees of $\tilde\Theta(\sqrt{m/T})$ where $m$ depends on the causal model and may be much smaller than $N$.
In contrast, existing best arm identification algorithms suffer $\Omega(\sqrt{N/T})$ simple regret.
This shows theoretically the value of our framework over the traditional bandit problem. 
Experiments in \S\ref{sec:experiments} further demonstrate the value of causal models in this framework.
\todom{Mention cumulative regret?}

In the general casual model interventions and observations may have a complex relationship. 
In \S\ref{sec:simple-regret-general} we propose an algorithm (Algorithm~\ref{alg:alloc}) inspired by importance-sampling that a) enjoys sub-linear regret equivalent to the optimal rate in the parallel bandit setting and b) captures many of the intricacies of sharing information in a causal graph in the general case. 
As in the parallel bandit case, the regret guarantee is often significantly better than $O(\sqrt{N/T})$ and never worse, where $N$ is the number of available interventions.
